BOLD Parameter Estimation Using Particle Filters

### 2010-07-11 doc/thesis/results.tex
simulated single voxel analysis done except for a little more discussion

\chapter{Results}

\section{Simulation}
I performed two types of simulations. First, I simulated a single BOLD time-series based
on a randomly selected set of model parameters. Second I used a modified version of the
FSL tool possum to generate a noisy FMRI image.

\subsection{Single Timeseries}
This process was relatively straight forward,
given the state-space equations for the BOLD signal. After a "true" signal was generated,
I then added a identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
process with Gaussian I.I.D. steps. Finally a carrier level was added, since BOLD is typically
measured as a \% difference from the
base level. Adding a carrier level meant that the exact same algorithm which was used for 
real data could be used for the simulated data.

@@

\subsubsection{Low Noise Simulation}
For the low noise case, the ten realization are shown in \autoref{fig:LowNoiseRealization}.
\begin{figure}
\label{fig:LowNoiseRealization}
\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/realization_lownoise}
\caption{Test Signals with low noise compared to the clean signal}
\end{figure}

@@

It becomes clear now the reason for the term particle filter. In several locations the output
of the particle filter looks like a filtered version of the input. For instance toward the
end, the estimates stay flat in spite of the preprocessed data drifting off a bit. By
this point, the algorithm had converged sufficiently to know that that wasn't possible.
A similar circumstance occurs at 100 seconds in. In general the fits are good, although because
of the slightly higher signal levels in the preprocessed signals, the fits tend to overestimate
the peak a bit. The final parameter sets are shown in \autoref{tab:LowNoiseParams}
\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
\hline 
$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & $\sqrt{MSE}$ (Res.) &$\sqrt{MSE}$\\
\hline 
\rowcolor[gray]{.8}
1.45 & .3 & .47 & .044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
\hline 
\hline 
1.22214 & 0.3449 & 0.33462 & 0.07138 & 1.60446 & 2.27529 & 1.59454 & 5.10189  &  0.00321067  & 0.00987647  \\
1.37493 & 0.33183 & 0.36296 & 0.07327 & 1.64076 & 2.10299 & 1.57627 & 5.11867 &  0.00305475  & 0.00993227  \\
1.16604 & 0.32205 & 0.34057 & 0.08217 & 1.64768 & 2.35351 & 1.24515 & 5.16722 &  0.00328932  & 0.00967961  \\
1.23176 & 0.32707 & 0.34028 & 0.07961 & 1.62698 & 2.18519 & 1.30326 & 5.04394 &  0.00284719  & 0.00912005  \\
1.1832 & 0.31787 & 0.34718 & 0.0821 & 1.54961 & 2.29115 & 1.27817 & 5.02396   &  0.00300634  & 0.00971258  \\
1.1424 & 0.33395 & 0.34725 & 0.07366 & 1.62208 & 2.29084 & 1.4025 & 5.05531   &  0.00283287  & 0.00948483  \\
1.30041 & 0.3596 & 0.35643 & 0.07679 & 1.56406 & 2.13234 & 1.60338 & 4.99681  &  0.00302802  & 0.01021885  \\
1.24008 & 0.34601 & 0.33978 & 0.08903 & 1.64989 & 2.23655 & 1.29004 & 5.12651 &  0.00304378  & 0.01007964  \\
1.1709 & 0.32739 & 0.34644 & 0.08255 & 1.53734 & 2.28262 & 1.37828 & 4.99087  &  0.00334488  & 0.01032886  \\
1.18967 & 0.34344 & 0.33554 & 0.07976 & 1.53582 & 2.30746 & 1.42774 & 5.03295 &  0.00317542  & 0.01001503  \\
1.184 & 0.34053 & 0.35017 & 0.08917 & 1.61025 & 2.27926 & 1.16448 & 5.07352   &  0.00288855  & 0.00950536  \\
\hline                                                                           
1.21868 & 0.33588 & 0.34557 & 0.07995 & 1.59899 & 2.24884 & 1.38762 & 5.06651 & 0.00306562     & 0.00981396 \\
\hline 
\end{tabular}
\caption{Estimated Parameters on 10 different runs with low noise. First row is the true parameters,
last is mean over the 10 runs. The $\sqrt{MSE}$ (Res.) is the square root of the mean square of the
residuals. Essentially this is the is the MSE between the estimated signal and the noisy signal which 
was available to the particle filter. Square root of MSE is the actual \emph{error}, with the true signal,
which of course was not available to the particle filter.}
\label{tab:LowNoiseResults} 
\end{table}

@@

\subsubsection{High Noise Simulation}
For the high noise simulation, the exact same procedure was performed again, but with 
measurement error and drift standard
deviations an order of magnitude higher. The noisy signals versus the base signal are shown
in \autoref{fig:HighNoiseRealization} and the preprocessed signals which the particle
filter attempt to fit are in \autoref{fig:PreprocessedHighNoise}.
\begin{figure}[H]
\label{fig:LowNoiseRealization}
\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/realization_highnoise}
\caption{Test Signals with high noise compared to the clean signal, $\sigma_x = .01, \sigma_y=.005$}
\end{figure}
\begin{figure}[H]
\label{fig:PreprocessedHighNoise}
\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/preprocessed_highnoise}
\caption{A comparison of the preprocessed signals for the high noise case.}
\end{figure}

@@

\begin{figure}[H]
\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
{\includegraphics[trim=7cm 3cm 7cm 4cm, width=15cm]{images/highnoise_run5_1}}\\
\end{figure}
\begin{figure}[H]
\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
{\includegraphics[trim=7cm 3cm 7cm 1cm, width=15cm]{images/highnoise_run5_2}}\\
\end{figure}
\begin{figure}[H]
\subfigure[$Q$, $S$, $F$, $BOLD$ ]
{\includegraphics[trim=7cm 3cm 7cm 2cm, width=15cm]{images/highnoise_run5_3}}
\label{fig:ConvergenceRuns1}
\caption{Converging histogram for parameters during run 2, as in \autoref{fig:NoiseComparisonJustTwo}.}
\end{figure}

@@

\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
\hline 
$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & $\sqrt{MSE}$ (Res.) & $\sqrt{MSE}$\\
\hline 
\rowcolor[gray]{.8}
1.45 & .3 & .47 & .044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
\hline 
\hline 
1.19002 & 0.23495 & 0.42228 & 0.128 & 1.01468 & 2.47795 & 1.11677 & 4.68265   &0.0140648  &0.01572919   \\
0.9721 & 0.21902 & 0.30505 & 0.061 & 0.57801 & 1.99596 & 3.46135 & 3.54608    &0.0137314  &0.01377685   \\
1.57947 & 0.14153 & 0.33798 & 0.1079 & 0.5843 & 2.12475 & 1.78343 & 4.28852   &0.01275372 &0.01577449   \\
1.10937 & 0.2374 & 0.53491 & 0.0351 & 1.21862 & 3.07365 & 2.35039 & 5.40164   &0.0167258  &0.0115433    \\
1.10712 & 0.27535 & 0.33651 & 0.03161 & 1.50567 & 2.65181 & 4.19099 & 5.2646  &0.01369793 &0.01221607   \\
0.58026 & 0.47931 & 0.41354 & 0.11894 & 0.97563 & 3.69018 & 1.00084 & 5.24607 &0.01149511 &0.01315637   \\
\rowcolor[rgb]{.9,.5,.5}
1.29515 & 0.25957 & 0.27559 & 0.25952 & 1.70265 & 2.8458 & 0.66172 & 5.84361  &0.01555035 &0.01789781   \\
\rowcolor[rgb]{.5,.5,.9}
1.5185 & 0.21987 & 0.28348 & 0.07417 & 0.88822 & 3.07706 & 1.73929 & 5.48378  &0.01205351 &0.01246202   \\
0.68736 & 0.3283 & 0.39786 & 0.15614 & 1.07781 & 3.1158 & 0.66432 & 4.88097   &0.01510364 &0.01257713   \\
1.01703 & 0.28497 & 0.34741 & 0.05672 & 1.58774 & 2.65157 & 2.28519 & 5.25634 &0.012493   &0.01343459   \\
0.99247 & 0.29795 & 0.32207 & 0.20943 & 0.42757 & 2.21081 & 1.01674 & 3.63085 &0.01216522 &0.01505545   \\
\hline                                                                          
1.09535 & 0.27075 & 0.36152 & 0.11259 & 1.05099 & 2.71958 & 1.84282 & 4.86592 &0.01362132  &0.01396575\\
\hline 
\end{tabular}
\caption{Estimated Parameters on 10 different runs with high noise. First row is the true parameters.
Note also that the blue row is Run 1 and the red row is Run 2, as used named this section}
\label{tab:HighNoiseResults} 
\end{table}

a series of tests to determine the convergence rate of the particle filter, the number
of particles that were required, how weighting functions compared, how different de-trending
methods compared with each other and, finally the variance of the result. By running the exact
time-series with different noise realizations, it was possible to determine the model variance.
As the reader may know, the error of an estimator may be calculated as:
\begin{equation}
MSE(\Theta) = Var(\Theta) + Bias(\Theta)^2
\end{equation}
The variance is an expression of how much the result would change for different noise realizations,
whereas the bias is an expression of how well the model matches the true underlying model. In
this case, because the same model is being used in the particle filter and underlying simulation,
the bias is actually zero. Obviously when this is calculated using \emph{real} data with an unknown
underlying state space equation, there will be some amount of bias error, but assuming that the
noise is similar to the noise used in these tests, the model variance will actually be about the
same. Thus calculating the model variance is extremely helpful in calculating how well determined
our model is, and how consistent it will be for real data. A single timeseries, as opposed to the
thousands present in a real image, makes it easier to 
compare the output with the ground truth, with various parameters.

Second I used a modified version of the FSL tool 
POSSUM to generate an entire FMRI image from a parameter map. The parameter map was generated
by taking an existing activation map and assigning discrete parameter sets to each region.
The result was a four dimensional (length x width
x height x parameter) image with spatially varying parameters. Possum was then modified
to take a parameter map and generate activation levels depending on the parameters at that
point. The patch for POSSUM will be made available. As an unfortunate side effect of 
not using Possums' original activation scale, I manually added 750 to the total level of
simulated Possum images. This is because the BOLD \% difference levels were in the range
of 50 - 100\% from the base, about 5 times as large as they should have been. Ultimately
this should not have an effect on the parameters (other than perhaps $\epsilon$ and $V_0$). 
For each time-series in the simulated FMRI image, the final \emph{static} parameters were saved
into a parameter map. This parameter map may then be compared to the map used to generate the 
simulated data; additionally a new simulation using the calculated parameters may also be 
generated to test the difference in activation levels between the real parameters and the
estimated ones. Since it is clear that the parameters are not fully orthogonal 
(\cite{Deneux2006}), its possible that two sets of parameters are functionally equivalent,
but have different parameters. This way, an absolute 
quantitative difference between the two parameter sets may be found.

>>>>

\chapter{Simulation}

\section{Single Timeseries}
I performed two types of simulations. First, I simulated a single BOLD time-series based
on a randomly selected set of model parameters. Second I used a modified version of the
FSL tool possum to generate a noisy FMRI image.

This process was relatively straight forward,
given the state-space equations for the BOLD signal. After a "true" signal was generated,
I then added a identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
process with Gaussian I.I.D. steps. Finally a carrier level was added, since BOLD is typically
measured as a \% difference from the
base level. Adding a carrier level meant that the exact same algorithm which was used for 
real data could be used for the simulated data.

@@

\subsection{Low Noise}
For the low noise case, the ten realization are shown in \autoref{fig:LowNoiseRealization}.
\begin{figure}
\label{fig:LowNoiseRealization}
\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/realization_lownoise}
\caption{Test Signals with low noise compared to the clean signal}
\end{figure}

@@

It becomes clear now the reason for the term particle filter. In several locations the output
of the particle filter looks like a filtered version of the input. For instance toward the
end, the estimates stay flat in spite of the preprocessed data drifting off a bit. By
this point, the algorithm had converged sufficiently to know that that wasn't possible.
A similar circumstance occurs at 100 seconds in. In general the fits are good, although because
of the slightly higher signal levels in the preprocessed signals, the fits tend to overestimate
the peak a bit. The final parameter sets are shown in \autoref{tab:LowNoiseParams}
\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
\hline 
$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & $\sqrt{MSE}$ (Res.) &$\sqrt{MSE}$\\
\hline 
\rowcolor[gray]{.8}
1.45 & .3 & .47 & .044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
\hline 
\hline 
1.22214 & 0.3449 & 0.33462 & 0.07138 & 1.60446 & 2.27529 & 1.59454 & 5.10189  &  0.00321067  & 0.00987647  \\
1.37493 & 0.33183 & 0.36296 & 0.07327 & 1.64076 & 2.10299 & 1.57627 & 5.11867 &  0.00305475  & 0.00993227  \\
1.16604 & 0.32205 & 0.34057 & 0.08217 & 1.64768 & 2.35351 & 1.24515 & 5.16722 &  0.00328932  & 0.00967961  \\
1.23176 & 0.32707 & 0.34028 & 0.07961 & 1.62698 & 2.18519 & 1.30326 & 5.04394 &  0.00284719  & 0.00912005  \\
1.1832 & 0.31787 & 0.34718 & 0.0821 & 1.54961 & 2.29115 & 1.27817 & 5.02396   &  0.00300634  & 0.00971258  \\
1.1424 & 0.33395 & 0.34725 & 0.07366 & 1.62208 & 2.29084 & 1.4025 & 5.05531   &  0.00283287  & 0.00948483  \\
1.30041 & 0.3596 & 0.35643 & 0.07679 & 1.56406 & 2.13234 & 1.60338 & 4.99681  &  0.00302802  & 0.01021885  \\
1.24008 & 0.34601 & 0.33978 & 0.08903 & 1.64989 & 2.23655 & 1.29004 & 5.12651 &  0.00304378  & 0.01007964  \\
1.1709 & 0.32739 & 0.34644 & 0.08255 & 1.53734 & 2.28262 & 1.37828 & 4.99087  &  0.00334488  & 0.01032886  \\
1.18967 & 0.34344 & 0.33554 & 0.07976 & 1.53582 & 2.30746 & 1.42774 & 5.03295 &  0.00317542  & 0.01001503  \\
1.184 & 0.34053 & 0.35017 & 0.08917 & 1.61025 & 2.27926 & 1.16448 & 5.07352   &  0.00288855  & 0.00950536  \\
\hline                                                                           
1.21868 & 0.33588 & 0.34557 & 0.07995 & 1.59899 & 2.24884 & 1.38762 & 5.06651 & 0.00306562     & 0.00981396 \\
\hline 
\end{tabular}
\caption{Estimated Parameters on 11 different runs with low noise. First row is the true parameters,
last is mean over the 10 runs. The $\sqrt{MSE}$ (Res.) is the square root of the mean square of the
residuals. Essentially this is the is the MSE between the estimated signal and the noisy signal which 
was available to the particle filter. Square root of MSE is the actual \emph{error}, with the true signal,
which of course was not available to the particle filter.}
\label{tab:LowNoiseResults} 
\end{table}

@@

\subsection{High Noise}
\label{sec:HighNoise}
For the high noise simulation, the exact same procedure was performed again, but with 
measurement error and drift standard
deviations an order of magnitude higher. The noisy signals versus the base signal are shown
in \autoref{fig:HighNoiseRealization} and the preprocessed signals which the particle
filter attempt to fit are in \autoref{fig:PreprocessedHighNoise}.
\begin{figure}[H]
\label{fig:LowNoiseRealization}
\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/realization_highnoise}
\caption{Test Signals with high noise compared to the clean signal, $\sigma_x = .01, \sigma_y=.005$}
\end{figure}
\begin{figure}[H]
\label{fig:PreprocessedHighNoise}
\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/preprocessed_highnoise}
\caption{A comparison of the preprocessed signals for the high noise case.}
\end{figure}

@@

\begin{figure}[H]
\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
{\includegraphics[trim=7cm 3cm 7cm 4cm, width=15cm]{images/highnoise_run5_1}}\\
\end{figure}
\begin{figure}[H]
\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
{\includegraphics[trim=7cm 3cm 7cm 1cm, width=15cm]{images/highnoise_run5_2}}\\
\end{figure}
\begin{figure}[H]
\subfigure[$Q$, $S$, $F$, $BOLD$ ]
{\includegraphics[trim=7cm 3cm 7cm 2cm, width=15cm]{images/highnoise_run5_3}}
\label{fig:ConvergenceRuns1}
\caption{Converging histogram for parameters during run 1, as in \autoref{fig:NoiseComparisonJustTwo}.}
\end{figure}

@@

\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
\hline 
$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & $\sqrt{MSE}$ (Res.) & $\sqrt{MSE}$\\
\hline 
\rowcolor[gray]{.8}
1.45 & .3 & .47 & .044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
\hline 
\hline 
1.19002 & 0.23495 & 0.42228 & 0.128 & 1.01468 & 2.47795 & 1.11677 & 4.68265   &0.0140648  &0.01572919   \\
0.9721 & 0.21902 & 0.30505 & 0.061 & 0.57801 & 1.99596 & 3.46135 & 3.54608    &0.0137314  &0.01377685   \\
1.57947 & 0.14153 & 0.33798 & 0.1079 & 0.5843 & 2.12475 & 1.78343 & 4.28852   &0.01275372 &0.01577449   \\
1.10937 & 0.2374 & 0.53491 & 0.0351 & 1.21862 & 3.07365 & 2.35039 & 5.40164   &0.0167258  &0.0115433    \\
1.10712 & 0.27535 & 0.33651 & 0.03161 & 1.50567 & 2.65181 & 4.19099 & 5.2646  &0.01369793 &0.01221607   \\
0.58026 & 0.47931 & 0.41354 & 0.11894 & 0.97563 & 3.69018 & 1.00084 & 5.24607 &0.01149511 &0.01315637   \\
\rowcolor[rgb]{.9,.5,.5}
1.29515 & 0.25957 & 0.27559 & 0.25952 & 1.70265 & 2.8458 & 0.66172 & 5.84361  &0.01555035 &0.01789781   \\
\rowcolor[rgb]{.5,.5,.9}
1.5185 & 0.21987 & 0.28348 & 0.07417 & 0.88822 & 3.07706 & 1.73929 & 5.48378  &0.01205351 &0.01246202   \\
0.68736 & 0.3283 & 0.39786 & 0.15614 & 1.07781 & 3.1158 & 0.66432 & 4.88097   &0.01510364 &0.01257713   \\
1.01703 & 0.28497 & 0.34741 & 0.05672 & 1.58774 & 2.65157 & 2.28519 & 5.25634 &0.012493   &0.01343459   \\
0.99247 & 0.29795 & 0.32207 & 0.20943 & 0.42757 & 2.21081 & 1.01674 & 3.63085 &0.01216522 &0.01505545   \\
\hline                                                                          
1.09535 & 0.27075 & 0.36152 & 0.11259 & 1.05099 & 2.71958 & 1.84282 & 4.86592 &0.01362132  &0.01396575\\
\hline 
\end{tabular}
\caption{Estimated Parameters on 10 different runs with high noise. First row is the true parameters.
Note also that the red row is Run 1 and the blue row is Run 2, as used named this section}
\label{tab:HighNoiseResults} 
\end{table}

\subsection{Pure-Noise, low magnitude}
The two final single-voxel test, forces the particle filter to attempt to "learn" a noise-only
timeseries. The noise used will be the same as that from the \autoref{sec:HighNoise},
$\sigma_x = .01, \sigma_y = .005$. The parameters will be set the same as well, but the
stimulus/input will be set to 0 the entire time. For the particle filter, on the other hand,
the stimulus will be left as in the previous two sections. The questions is if the particle
filter is able to appropriately differentiate from a signal with noise. The noise timeseries
are shown in \autoref{fig:NoiseOnly}, the preprocessed versions are shown in \autoref{fig:NoiseOnlyPrep}.
\begin{figure}[H]
\label{fig:NoiseOnly}
\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/realization_noiseonly}
\caption{Time-series lacking any real signal. With, $\sigma_x = .01, \sigma_y=.005$}
\end{figure}
\begin{figure}[H]
\label{fig:PreprocessedNoiseOnly}
\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/preprocessed_noiseonly}
\caption{A comparison of the preprocessed signals for the signal-free case.}
\end{figure}

\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c |}
\hline 
$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  &  $\sqrt{MSE}$   \\
\hline 
1.0324 & 0.33211 & 0.34058 & 0.03012 & 1.40665 & 2.52079 & 0.5311 & \\
 0.98189 & 0.33047 & 0.3386 & 0.03014 & 1.45707 & 2.47232 & 0.45049 & 0.00158551 \\
 1.0429 & 0.33224 & 0.34124 & 0.02946 & 1.4618 & 2.49245 & 0.43012 &  0.00165055 \\
 1.02054 & 0.3321 & 0.33484 & 0.02586 & 1.45848 & 2.48741 & 0.4193 &  0.00150604 \\
 1.0565 & 0.33405 & 0.33758 & 0.02791 & 1.43784 & 2.52545 & 0.47517 & 0.00151513 \\
 1.01867 & 0.33528 & 0.33918 & 0.02782 & 1.48345 & 2.49605 & 0.44209 &0.00155645 \\
 1.051 & 0.33038 & 0.33837 & 0.02985 & 1.47651 & 2.48621 & 0.42719 &  0.00158526 \\
 1.00281 & 0.32929 & 0.33988 & 0.0298 & 1.43519 & 2.49256 & 0.48899 & 0.00164299 \\
 1.00893 & 0.33273 & 0.33982 & 0.0289 & 1.42903 & 2.49754 & 0.45688 & 0.0016793  \\
 1.01289 & 0.33275 & 0.3376 & 0.02997 & 1.41188 & 2.49881 & 0.50628 & 0.00182622 \\
 1.10247 & 0.33371 & 0.3419 & 0.02939 & 1.43774 & 2.53384 & 0.44079 & 0.00195009 \\
\hline                                                                
1.03009 & 0.33228 & 0.33905 & 0.02902 & 1.44506 & 2.50031 & 0.46076 & 0.00165129 \\
\hline 
\end{tabular}
\caption{Estimated Parameters on 10 different runs with low noise. }
\label{tab:NoiseOnlyResults} 
\end{table}

The line fits are shown in \autoref{fig:fits_noiseonly}. Now the
low MSE and reasonableness of the parameters is cause for concern. Additionally, the precision
with which the convergence occurred is enviable, compared to the actual signals.
However, the impulse response, which has less importance in nonlinear systems, could
be usefull for separating out good voxels from bad. Note that in \autoref{fig:PreprocessedNoiseOnly}
peaks at well below $.005$, the standard deviation of the weighting function. Thus, the reason
the fits are in such agreement is that no convergence occurred at all, as 
further shown by \autoref{JustNoiseConvergence}. Thus one more test is necessary on
a signal with enough noise to actually be considered a signal. But first,
there is still the matter of identifying this noise-only voxel. Clearly MSE and
the parameters themselves don't have the information to differentiate in this case.
Thus the answer falls on either the final estimated variance (which will be high
because no particles strayed far from the center) or the impulse response.
I prefer the impulse response because it doesn't depend on something that 
may be less than dependable, namelay the rate of convergence.

\begin{figure}[H]
\label{fig:fits_noiseonly}
\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/fits_noiseonly}
\caption{Fits to the non-active, low noise signal. Note that the line is thick because all
the fits overlap. This is all 1 fitted lines.}
\end{figure}

\begin{figure}[H]
\label{fig:justnoise_fit_0}
\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/justnoise_fit_0}
\caption{Fit from a single particle filter run, with the noise input. }
\end{figure} %uses allnoise/ALLNOISE-0-w0

\begin{figure}[H]
\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
{\includegraphics[trim=7cm 3cm 7cm 4cm, width=15cm]{images/justnoise_hist_1}}\\
\end{figure}
\begin{figure}[H]
\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
{\includegraphics[trim=7cm 3cm 7cm 1cm, width=15cm]{images/justnoise_hist_2}}\\
\end{figure}
\begin{figure}[H]
\subfigure[$Q$, $S$, $F$, $BOLD$ ]
{\includegraphics[trim=7cm 3cm 7cm 2cm, width=15cm]{images/justnoise_hist_3}}
\label{fig:JustNoiseConvergence}
\caption{Converging histogram for parameters when the signal consists purely of low level noise. 
Same run as \autoref{fig:justnoise_fit_0}}
\end{figure}

To give a frame of reference to the impulse response level, the impulse response of every
parameter estimated, includeing ones from \autoref{sec:BigNoise}, are shown in \autoref{tab:ImpulseResponses}.

\subsection{Pure Noise, High Magnitude}
\label{sec:PureNoiseHighMag}

\begin{figure}[H]
\label{fig:fits_noiseonly_high}
\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/fits_noiseonly_high}
\caption{Fits to the non-active, high noise signal. Note that the line is thick because all
the fits overlap. This is all 11 fitted lines.}
\end{figure}

\begin{figure}[H]
\label{fig:justbignoise_fit_0}
\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/justbignoise_fit_0}
\caption{Fit from a single particle filter run, with the noise input. }
\end{figure} %uses allnoise/ALLNOISE-0-w0

\begin{figure}[H]
\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
{\includegraphics[trim=7cm 3cm 7cm 4cm, width=15cm]{images/justbignoise_1}}\\
\end{figure}
\begin{figure}[H]
\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
{\includegraphics[trim=7cm 3cm 7cm 1cm, width=15cm]{images/justbignoise_2}}\\
\end{figure}
\begin{figure}[H]
\subfigure[$Q$, $S$, $F$, $BOLD$ ]
{\includegraphics[trim=7cm 3cm 7cm 2cm, width=15cm]{images/justbignoise_3}}
\label{fig:JustNoiseConvergenceLarge}
\caption{Converging histogram for parameters when the signal consists purely of noise, with peaks comparable  to 
the convential BOLD signal. ($\sigma_y = .01, \sigma_x = .005$). Same run as fit in \autoref{fig:justbignoise_fit_0}}
\end{figure}

\subsection{Single Voxel Review}
\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c |}
\hline 
& \multicolumn{2}{|c|}{Signal} & \multicolumn{2}{|c|}{No Signal}\\
\hline
Run  & Low Noise & High Noise & $\sigma_y=.001, \sigma_x=.0005$ & $\sigma_y=.01, \sigma_x=.005$\\
\hline 
\hline       
1 &    0.0115482 &       0.015527 &        0.00138579 &      0.00843435   \\
2 &    0.0116813 &       0.0126847 &       0.00152434 &      0.00495585   \\
3 &    0.011314 &        0.0123151 &       0.0014101 &       0.00667397  \\
4 &    0.0107892 &       0.0114702 &       0.00119941 &      0.0038459    \\
5 &    0.0112809 &       0.0143503 &       0.00145239 &      0.00811513   \\
6 &    0.0112021 &       0.0113265 &       0.00137792 &      0.00662327   \\
7 &    0.011943 &        0.0192599 &       0.00141883 &      0.00734546   \\
8 &    0.0118222 &       0.010144 &        0.00162352 &      0.00817847   \\
9 &    0.0120212 &       0.0122466 &       0.00146133 &      0.00843033   \\
10 &   0.0115679 &       0.0156513 &       0.00165983 &      0.00819587   \\
11 &   0.0110983 &       0.0103732 &       0.001415 &        0.00886845  \\
\hline
Mean& 0.0114789417183 &  0.0132135264575 & 0.00144804221451 &0.00724245910533 \\
\hline
\end{tabular}
\caption{Estimated Impulse response on 11 different for each of the cases covered in this chapter.}
\label{tab:SignleVoxelActivationComparison} 
\end{table}

\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c |}
\hline 
& \multicolumn{2}{|c|}{Signal} & \multicolumn{2}{|c|}{No Signal}\\
\hline
Run  & Low Noise & High Noise & $\sigma_y=.001, \sigma_x=.0005$ & $\sigma_y=.01, \sigma_x=.005$\\
\hline 
\hline       
1 &   0.00321067 & 0.0140648  &   0.00166666 &  0.01576746 \\
2 &   0.00305475 & 0.0137314  &   0.00158551 &  0.0155104 \\
3 &   0.00328932 & 0.01275372  &  0.00165055 &  0.01395712 \\
4 &   0.00284719 & 0.0167258  &   0.00150604 &  0.01381387 \\
5 &   0.00300634 & 0.01369793  &  0.00151513 &  0.01435028 \\
6 &   0.00283287 & 0.01149511  &  0.00155645 &  0.01350408 \\
7 &   0.00302802 & 0.01555035  &  0.00158526 &  0.01300628 \\
8 &   0.00304378 & 0.01205351  &  0.00164299 &  0.01544561 \\
9 &   0.00334488 & 0.01510364  &  0.0016793 &   0.01484681 \\
10 &  0.00317542 & 0.012493  &    0.00182622 &  0.01312527 \\
11 &  0.00288855 & 0.01216522  &  0.00195009 &  0.01445691 \\
\hline
Mean& 0.00306562 & 0.01362132  &  0.00165129 &  0.01434401 \\
\hline
\end{tabular}
\caption{MSE with response to the preprocessed, noisy signal for 11 runs for each of the cases covered in this chapter.}
\label{tab:SingleVoxelMSEComparison} 
\end{table}

Comparing the results of \autoref{sec:HighNoise} and \autoref{sec:PureNoiseHighMag}, it is
clear that distinguishing between these cases is a difficult prospect. Although
the impulse response, which is determined by the peak response to a very short input,
is is slightly different, there is no reason to believe that a noise distribution
could not get a higher impulse response. The $\sqrt{MSE}$ is also not going to be
able to distinguish between the high-noise case and the no signal case. Therefore, there
is a noise ceiling above which it will be impossible to determine model validity. The fact
that the particle filter is able to get close to the "true" signal is sadly irrelevant.
Ultimately, a threshold well below those shown in the high noise case 
will have to be placed on the error, if any confidence is going to be placed on the results.

Each test took approximately 60 seconds on an 8 core Xeon (specifics?).

\section{Multi-voxel Simulation}
To test the usefullness of the particle filter for Second I used a modified version of the FSL tool 
POSSUM to generate an entire FMRI image from a parameter map. The parameter map was generated
by taking an existing activation map and assigning discrete parameter sets to each region.
The result was a four dimensional (length x width
x height x parameter) image with spatially varying parameters. Possum was then modified
to take a parameter map and generate activation levels depending on the parameters at that
point. The patch for POSSUM will be made available. As an unfortunate side effect of 
not using Possums' original activation scale, I manually added 750 to the total level of
simulated Possum images. This is because the BOLD \% difference levels were in the range
of 50 - 100\% from the base, about 5 times as large as they should have been. Ultimately
this should not have an effect on the parameters (other than perhaps $\epsilon$ and $V_0$). 
For each time-series in the simulated FMRI image, the final \emph{static} parameters were saved
into a parameter map. This parameter map may then be compared to the map used to generate the 
simulated data; additionally a new simulation using the calculated parameters may also be 
generated to test the difference in activation levels between the real parameters and the
estimated ones. Since it is clear that the parameters are not fully orthogonal 
(\cite{Deneux2006}), its possible that two sets of parameters are functionally equivalent,
but have different parameters. This way, an absolute 
quantitative difference between the two parameter sets may be found.
